Does Anybody Really Know What Time It Is?\\ Automating the Extraction of Date Scalars

### 2015-07-08 nlp.tex
Added NLP book citation
Added grammar definition

Pattern-recognition problems such as parsing date and time formats initiate from observations generated by some structured stochastic process. In other words, even if the initial higher-level production rule of the grammar is known (i.e. date, time or date-time), there could be several directions that the parser resolve to. For example, in a date string 5/6/2015, the pattern could either be M/d/yyyy or d/M/yyyy.

>>>>

While there are several functionally equivalent notations for representing CFG, we use the \textit{Backus-Naur Form} (BNF) for defining the grammar rules for \dateparse formats. In particular, we use the \textit{Extended Backus-Naur Form} (EBNF) as the notation is more compact and readable for frequently used constructions~\cite{Grune:1990}.

We define a grammar for identifying date-time strings based on other EBNF based date-time formats [cite] as a reference. A partial definition of the grammar is found below (see supplementary material for complete definition):

\begin{grammar}
<TimeGrammar> ::= <Hours> ':' <Minutes> ':' <Seconds> <AMPM>;\\

<DateGrammar> ::= <BigEndianDate> 
				\alt <MiddleEndianDate> 
				\alt <LittleEndianDate>;

<DateTimeGrammar>  ::= <DateGrammar> 
					\alt <TimeGrammar>;

<AMPM>  ::=  WORD('a.m.') 
			 \alt WORD('p.m.') ;

<Year> ::= <TwoYear> | <FourYear>;

<Month>  ::= <MonthFullForm> | <MonthAbbrForm> | <MonthLetterForm> | <MonthNumber>;

<BigEndianDate> ::= <Year> <Month>  <Day> ;

<MiddleEndianDate> ::= <Month> <Day> <Year>;

<LittleEndianDate> ::= <Day> <Month> <Year>;

\end{grammar}

\subsection{Translation of Grammar Output to ICU Format}

\subsection{Production Rule Constraints and Variants}

\subsection{Probabilistic Context-Free Grammar}
Pattern-recognition problems such as parsing date and time formats initiate from observations generated by some structured stochastic process. In other words, even if the initial higher-level production rule of the grammar is known (i.e. date, time or date-time), there could be several directions that the parser resolve to. For example, in a date string 5/6/2015, the pattern could either be M/d/yyyy or d/M/yyyy.
